\begin{abstract}
In November of 2007, the National Institute for Standards and Technology (NIST) announced the request for submission of candidate algorithms for the Secure Hash Algorithm 3 (SHA-3).
Because of recent attacks against the current standards, NIST ``decided that it is prudent to develop a new hash algorithm to augment and revise FIPS 180–2'' \cite{Sha3Request}.

As the second round of candidates were recently chosen in September, we decided to choose few of these second-round candidates--CubeHash \cite{CubeHash}, Skein \cite{Skein}, and ECHO \cite{ECHO}--and see if it is feasible to implement them in a GPU computing architecture.
In particular, we chose to implement them in Open Computing Language (OpenCL), which provides a framework for executing ``kernels'' on a variety of devices (e.g., CPUs, GPUs, etc.).

We found that while all the hash algorithms performed more poorly on the NVIDIA GPU than the CPU, the ECHO hash algorithm actually performed faster on the ATI GPU than it did on either the NVIDIA GPU or the CPU.
The ECHO algorithm on the ATI GPU even performed faster than the CubeHash algorithm on the CPU.
This is a very exciting result, showing that there is reason to beleive that GPUs can be a useful tool when performing cryptographic functions.
However, this result should be taken with a grain of salt - while the actual run times as reported by the OpenCL profile timer are faster for the ATI GPU than the CPU, this result runs counter to all of the other data points we collected.
That said, we were skeptical of the results at first, and reran the test suite 3 additional times, receiving the same results each time.
However, despite the runtimes being significantly faster on the GPU than the CPU, the overall runtime for the test suite when overhead is taken into account was about 5 times slower than the CPU version.

We also found that the Skein algorithm is by far the fastest (comparable to SHA1) when run only on a CPU - even with the OpenCL overhead it comes close to the theoretical speed claimed by the authors.
While the Skein algorithm is slower when run on the GPU than most of the other algorithms, for the average user the Skein algorith mis likely the best option - most people do not have a Radeon 4870 in their computer.

\end{abstract}
