\begin{thebibliography}{9}
  \bibitem{Sha3Request} ``Announcing Request for Candidate Algorithm Nominations for a New Cryptographic Hash Algorithm (SHAÐ3) Family.'' Federal Registrar, Vol. 72, No. 212 (November 2007), pp. 62212-62220. Available at \url{http://csrc.nist.gov/groups/ST/hash/documents/FR_Notice_Nov07.pdf}.
  \bibitem{Nostradamus} Marc Stevens, Arjen K. Lenstra, Benne de Weger. ``Predicting the winner of the 2008 US Presidential Elections using a Sony PlayStation 3.'' November 30, 2007. Available at \url{http://www.win.tue.nl/hashclash/Nostradamus/}
  \bibitem{Sha1Collisions} Xaoyun Wang, Yiqun Lisa Yin, Hongbo Yu. ``Finding Collisions in the Full SHA-1.'' Lecture Notes in Computer Science, Vol. 3621 (November 2005), pp. 17-36
  \bibitem{Sha3Zoo} ``The SHA-3 Zoo.'' Available at \url{http://ehash.iaik.tugraz.at/wiki/The_SHA-3_Zoo}
  \bibitem{Sha3Hardware} ``SHA-3 Hardware Implementations.'' Available at \url{http://ehash.iaik.tugraz.at/wiki/SHA-3_Hardware_Implementations}
  \bibitem{Westmere} ``Intel Nehalem (microarchitecture).'' Available at \url{http://en.wikipedia.org/wiki/Intel_Westmere}
  \bibitem{CubeHash} ``CubeHash: a simple hash function.'' Available at \url{http://cubehash.cr.yp.to/}
  \bibitem{CubeHash-spec} Bernstein, Daniel J. ``CubeHash specification (2.B.1).'' Available at \url{http://cubehash.cr.yp.to/submission/spec.pdf}.
  \bibitem{Skein} ``The Skein Hash Function Family.'' Available at \url{http://www.skein-hash.info/} and \url{http://www.schneier.com/skein.html}
  \bibitem{ECHO} ``ECHO Hash Function.'' Available at \url{http://crypto.rd.francetelecom.com/echo/}.
  \bibitem{Harrison} Owen Harrison and John Waldron. ``AES Encryption Implementation and Analysis on Commodity Graphics Processing Units.'' Available at \url{https://www.cs.tcd.ie/~harrisoo/publications/AES_On_GPU.pdf}.
  \bibitem{CodeBook} ``Block cipher modes of operation'' Available at \url{http://en.wikipedia.org/wiki/Block_cipher_modes_of_operation#Electronic_codebook_.28ECB.29}.
  \bibitem{Manavski} ``CUDA COMPATIBLE GPU AS AN EFFICIENT HARDWARE ACCELERATOR FOR AES CRYPTOGRAPHY.'' Available at \url{http://www.manavski.com/downloads/PID505889.pdf}
  \bibitem{SkeinSpeed} ``The Skein Hash Function Family.'' Available at \url{http://www.schneier.com/skein.html}
  \bibitem{Bernstein} Daniel J. Bernstein. ``CubeHash parameter tweak: 16 times faster.'' Available at \url{http://cubehash.cr.yp.to/submission/tweak.pdf}
\end{thebibliography}
